A Global Genome-wide Scan with Optimal Cutoff Mining for Emerging Biomarkers in Head and Neck Squamous Cell Carcinoma

### 2021-03-10 main.tex
ERROR free
Updates from Overleaf
truncated TCIA subsection

% I have a graphic abstract: graphic_abstract_pvalueTex.pdf
% National Cancer Institute. Changing the dressing on a patient’s neck. https://unsplash.com/photos/IrqLPsxemQ4?utm_source=unsplash&utm_medium=referral&utm_content=creditShareLink (1950).\cite{nci1950}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\begin{paracol}{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{section}{-1} %% Remove this when starting to work on the template.
%\section{How to Use this Template}

@@

%\subsubsection{Figure 1:h}
\begin{figure}[hp]
\centering
\includegraphics[width=14cm]{Figure_1_manuscript_workflow} % .PDF is better than .png
%, height=8cm
%\caption
\bcaption{A workflow of \acrshort{hnscc} biomarker discovery, step 1 (\textcolor{blue}{blue line}: main procedure) and step 2 (\textcolor{orange}{orange line}: analysis export).}
%Step 3 (purple line: dealing with surgical margin).
{The "main procedure" includes data retrieval from TCGA GDC data portal, data process with merging and cleaning (red rectangled),% should be modified aggressively according to different sources of database)
then performing the survival analyses. The Cutoff engine (cutofFinder\_func.HNSCC.R) might calculate all possible Kaplan-Meier \textit{P} value to find the optimal cutoff value of RNA-Seq for subsequent Cox modeling (a draft diagram shown on the upper right corner "HNSCC cohort", the serial cut for grouping patients with \textcolor{green}{low [green]} or \textcolor{red}{high [red]} expression of a specific gene, to yield a collection of \textit{P} values; please see Materials and Methods section for details). The step 2 "analysis export" performs dissecting and selection of candidate genes by Bonferroni adjusted \textit{P} value as well as a hazard ratio of Cox model, based on the results from the step 1.
(HNSCC: head and neck squamous cell carcinoma; TCGA: the Cancer Genome Atlas; RNA-Seq: RNA sequencing; GDC: Genomic Data Commons.)}
\label{fig:figure1}
\end{figure}

@@

\begin{paracol}{2}
\linenumbers
\switchcolumn

@@

\subsubsection{TCIA validation by deep learning}
%https://wiki.cancerimagingarchive.net/display/Public/TCGA-HNSC
%https://cancer.digitalslidearchive.org/#!/CDSA/hnsc/TCGA-BA-4076/TCGA-BA-4076-01A-01-BS1?slide=TCGA-BA-4076-01A-01-BS1.4aef481c-f634-4ffc-9fb3-609f59a8caf9.svs
TCGA Pathology Images => The Cancer Imaging Archive (TCIA) by Deep learning (R/Keras framework)
%數位模擬，能讓臨床醫師和研究人員更快速了解細胞成長後的狀況，而不是花費時間等待細胞的培養。
image-based phenotypes and their underlying genomic data 
%https://www.biorxiv.org/content/10.1101/2020.03.10.985887v1.full
%https://github.com/Serian1992/ImgBio.

%The Stanford Tissue Microarray Database (TMAD; x http://tma.stanford.edu) 
%https://tma.im/cgi-bin/ontologyBrowser.pl?anchor=C34447; there is few HNSCC slides

*** interpretable or explainable CNN
%Manning meap book: 不再是黑盒子
%https://www.manning.com/books/trust-in-machine-learning
Trust in Machine Learning

%*** 必備 explainable interpretable 
By saliency mapping: e.x. diabetes retinopathy, model focused on retinal aterry

%https://bmcmedgenomics.biomedcentral.com/articles/10.1186/s12920-019-0624-2
%#XAI
Interpretable deep neural network for cancer survival analysis by integrating genomic and clinical data

Jie Hao, 2019
By PyTorch

%https://www.deeplearningbook.org/
%https://www.cancer.gov/research/areas/diagnosis/artificial-intelligence
% NCI
Aiding the Genomic Characterization of Tumors: 
deep learning could predict commonly mutated genes and mRNA expression from the images
*** A deep learning model to predict RNA-Seq expression of tumours from whole slide images\cite{Schmauch2020}
Preprocessing of whole-slide images. The application of deep-learning algo- rithms to histological data is a challenging problem, particularly due to the high dimensionality of the data (up to 100,000 × 100,000 pixels for a single whole-slide image) and the small size of available datasets. We divided the whole-slide images into squares of 112 × 112 μm (224 × 224 pixels) called “tiles”, and used the Otsu algorithm62 (as implemented in python package skimage) to select only those containing tissue, excluding the white background. We sampled a maximum of 8000 such tiles from each slide. We then extracted 2048 features from those tiles with a 50-layer ResNet63 pretrained on the ImageNet dataset64 (using the Keras implementation), such that a slide could be represented as a 8000 × 2048 matrix.
For the first phase of this work (transcriptome prediction), we accelerated the training of our models through a simple preprocessing step inspired by simple linear iterative clustering (SLIC)65: we used the k-means algorithm (as implemented in python package libkmcuda) to create 100 clusters (supertiles) of tiles on the basis of tile location on the slide, and we averaged the features of the tiles within each cluster. The use of these supertiles reduces the dimensions of a slide to 100 × 2048. The model was first trained on this reduced dataset, with all the TCGA data. Then, for specific organs, fine-tuning was achieved with full-scale data from the organs concerned only.

%AI methods can also be used to identify specific gene mutations from tumor pathology images instead of using traditional genomic sequencing. For instance, NCI-funded researchers at New York University used deep learning (DL) to analyze pathology images of lung tumors obtained from The Cancer Genome Atlas. Not only could the DL method accurately distinguish between two of the most common lung cancer subtypes, adenocarcinoma and squamous cell carcinoma, it could predict commonly mutated genes from the images.

Understanding the Method Behind the Machine: interpredability

%One challenge of AI, and DL specifically, is the “black box” problem: not fully understanding what features of the data a computer has used in its decision-making process. For example, a DL algorithm that predicts the optimal treatment for a patient does not provide the reasoning it used to make that prediction. Additional efforts are needed to reveal how algorithms arrive at a decision or prediction so that the process becomes transparent to scientists and clinicians. Making these algorithms transparent could help researchers identify new biological features relevant to disease diagnosis or treatment.
***  unconscious biases:
Incorporating information about biological processes into the algorithm is likely to improve its accuracy and decrease dependence on large amounts of annotated data, which may not be available. One danger of the “black box” problem is that DL may inadvertently perpetuate existing unconscious biases. Researchers need to carefully consider how potential biases affect the data being used to develop a model, adopt practices to address and monitor those biases, and monitor performance and applicability of AI models.

%With increased investments, NCI’s efforts to realize AI’s potential will lead to more accurate and rapid diagnoses, improved clinical decision-making, and, ultimately, better health outcomes for patients with cancer and those at risk. Updated: August 31, 2020
%
TMA by deep learning (CNN) mRNA expression associated with pathological images
*** TCGA TCIA Pathology Images with deep learning: gene expression on pathological section, or digital tissue microarray, TMA
%recurrent neural network for NLP

%佩戴可監測心率、體溫、血壓及其他關鍵數據的裝置，醫師也因此能夠獲得更多資料以更深入瞭解患者的病情，不再依靠病人的回憶進行判斷，診斷的準確性也會隨之提升。

%在生物醫學領域，氨基酸或DNA等序列的使用可說是相當普遍，由於序列可視為一種具有隱性結構的語言，因此自然語言處理模型所採用的架構便可運用在理解並生成生物醫學領域的序列。

%the Rstudio Shiny app (https://shiny.rstudio.com) as a web-based tool (named "pvalueTex" ) packaged with our workflow in the future.
%with more TCGA features (holistic)

@@

We also aim to promote a reproducible bioinformatics\cite{Preeyanon2014}\cite{Kulkarni2018} workflow allowing successful repetition and extension of analyses based on all the TCGA or other in-house cancer datasets.

%Dali Lama: Wish and bless the whole world %慈悲心 利他心
% 多了解身心靈的關係，減少癌症對健康的損害

>>>>

% I have a graphic abstract: graphic_abstract_pvalueTex.pdf
% National Cancer Institute. Changing the dressing on a patient’s neck. https://unsplash.com/photos/IrqLPsxemQ4?utm_source=unsplash&utm_medium=referral&utm_content=creditShareLink (1950).\cite{nci1950}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\globalcounter* % for paracol

\begin{document}
%\linenumbers
%\begin{paracol}{2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{section}{-1} %% Remove this when starting to work on the template.
%\section{How to Use this Template}

@@

%\subsubsection{Figure 1:h}
\begin{figure}[hp]
\centering
%\widefigure
\includegraphics[width=14cm]{Figure_1_manuscript_workflow} % .PDF is better than .png
%, height=8cm
%\caption
\bcaption{A workflow of \acrshort{hnscc} biomarker discovery, step 1 (\textcolor{blue}{blue line}: main procedure) and step 2 (\textcolor{orange}{orange line}: analysis export).}
%Step 3 (purple line: dealing with surgical margin).
{The "main procedure" includes data retrieval from TCGA GDC data portal, data process with merging and cleaning (red rectangled),% should be modified aggressively according to different sources of database)
then performing the survival analyses. The Cutoff engine (cutofFinder\_func.HNSCC.R) might calculate all possible Kaplan-Meier \textit{P} value to find the optimal cutoff value of RNA-Seq for subsequent Cox modeling (a draft diagram shown on the upper right corner "HNSCC cohort", the serial cut for grouping patients with \textcolor{green}{low [green]} or \textcolor{red}{high [red]} expression of a specific gene, to yield a collection of \textit{P} values; please see Materials and Methods section for details). The step 2 "analysis export" performs dissecting and selection of candidate genes by Bonferroni adjusted \textit{P} value as well as a hazard ratio of Cox model, based on the results from the step 1.
(HNSCC: head and neck squamous cell carcinoma; TCGA: the Cancer Genome Atlas; RNA-Seq: RNA sequencing; GDC: Genomic Data Commons.)}
\label{fig:figure1}
\end{figure}

@@

\begin{paracol}{2}
\linenumbers
\switchcolumn[1]

@@

@@

We also aim to promote a reproducible bioinformatics\cite{Preeyanon2014}\cite{Kulkarni2018} workflow allowing successful repetition and extension of analyses based on all the TCGA or other in-house cancer datasets.
%the Rstudio Shiny app (https://shiny.rstudio.com) as a web-based tool (named "pvalueTex" ) packaged with our workflow in the future.
%with more TCGA features (holistic)
%Dali Lama: Wish and bless the whole world %慈悲心 利他心
% 多了解身心靈的關係，減少癌症對健康的損害
